---
layout: page
link: https://www.brianomeara.info/
slug: home
title: Home
output:
  html_document:
    includes:
       in_header: GA_Script.html
---

```{r render_md, include=FALSE, echo=FALSE, warning=FALSE}
#library(remotes)
#remotes::install_github("bomeara/cv")

```

_Most of this website is written for a specialist scientific audience; folks who aren't scientists are very welcome to look around, but [this page](general.qmd) might be a more accessible introduction. For a compressed version of information about me, please see my [CV](cv.qmd)._

_Also note that this is my personal website, and any views expressed here are my own and not necessarily those of current, past, or future employers. For my official webpage, go to <https://eeb.utk.edu/people/brian-omeara/>_ .

My research involves the development and application of phylogenetic tools to address evolutionary and ecological questions.

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, eval=FALSE}
library(scholar)
library(plotly)

id <- 'vpjEkQwAAAAJ'
pubs <- scholar::get_publications(id)
pubs <- pubs[!is.na(pubs$year),]
pubs <- pubs[order(pubs$year, decreasing=TRUE),]

preprints <- pubs[grepl("xiv", pubs$journal),]
not_preprints <- pubs[!grepl("xiv", pubs$journal),]
not_preprints <- not_preprints[!grepl("xiv", not_preprints$title),]
not_preprints <- not_preprints[nchar(not_preprints$journal) > 0,]

not_preprints$description <- paste0(not_preprints$title, " (in ", not_preprints$journal, ")")
not_preprints$cites_plus1 <- not_preprints$cites+1
not_preprints$log1pcites <- log1p(not_preprints$cites)
fig <- plot_ly(data=not_preprints, x=~year, y=~log1pcites, type='scatter', mode='markers', text=~description, hoverinfo='text') %>% config(displayModeBar = FALSE) %>% 
  layout(title='Recent publications', xaxis=list(title='Year', range=c(2012,1+as.integer(format(Sys.Date(), "%Y")))), yaxis=list(title='Citations', 
  zerolinecolor = '#ffff',
	zerolinewidth = 20,
    gridcolor = 'ffff',
	tickvals = as.list(log1p(c(0,1, 10, 100, 500, 1000))),
	ticktext = as.list(c(0,1, 10, 100, 500, 1000)))
	)
```


```{r, echo=FALSE, warning=FALSE, eval=FALSE}
fig
```

They are usually generated by a direct research need: how can we tell whether this group is evolving at a different rate? How can we choose between phylogeographic models without limiting ourselves to a pre-selected small set? Is there hidden variation in states that lets some herbaceous plants retain the ability to make wood while others have lost this ability? By developing techniques to address these questions, we both solve the original question and enable other biologists to use these new techniques to answer more questions. Broadly, the areas covered include trait evolution, species delimitation, phylogeography, dating trees, and more work in progress.


## Research


I understand evolutionary biology by creating and applying new approaches. A major focus is on problem-solving: "we want to understand this, but we don't have the tools to do it (at least not well enough)." Like much of science, this is often done in collaboration. Some recent examples of the work:

### Rates of evolution

Evolution does not proceed at a constant rate or direction. We can use this to test hypotheses about what causes this heterogeneity: does eating this prey correlate with faster specialization of the jaw? Does a change of habitat lead to a different body size optimum? 

* "A novel method for jointly modeling the evolution of discrete and continuous traits." [Paper](https://doi.org/10.1093/evolut/qpad002) [Software](http://thej022214.github.io/OUwie/)
* "A Spatially Explicit Model of Stabilizing Selection for Improving Phylogenetic Inference ." [Paper](https://doi.org/10.1093/molbev/msaa318) [Software](https://github.com/bomeara/selac)


### Diversification

Why are there so many species? What traits affect rates of species persistance and change through time? 

* "A flexible method for estimating tip diversification rates across a range of speciation and extinction scenarios." [Paper](https://doi.org/10.1111/evo.14517) [Software](https://speciationextinction.info/)
* "Detecting hidden diversification shifts in models of trait-dependent speciation and extinction." [Paper](https://doi.org/10.1093/sysbio/syw022) [Software](https://speciationextinction.info/)

### Other interesting questions

* "dentist: Computing uncertainty by sampling points around maximum likelihood estimates." [Preprint](https://doi.org/10.1101/2023.01.10.523430) [Software](https://bomeara.github.io/dentist/)
* "Retiring “Cradles” and “Museums” of Biodiversity." [Paper](https://doi.org/10.1086/717412)
* "The evolution of a tropical biodiversity hotspot." [Paper](https://doi.org/10.1126/science.aaz6970)
* "The ecological and genomic basis of explosive adaptive radiation." [Paper](https://doi.org/10.1038/s41586-020-2652-7)
* "PHRAPL: phylogeographic inference using approximate likelihoods." [Paper](https://doi.org/10.1093/sysbio/syx001) [Software](http://www.phrapl.org/)

## Teaching

I've taught everything from 200-person introductory biology courses to small graduate reading groups. Every Fall I teach a macroevolution course for upper level undergraduate students and graduate students. I also taught an [open graduate course in phylogenetic methods](https://phylometh.info) every spring for years as part of an NSF CAREER grant. I teach our graduate core course in evolution. I also organize and/or teach at a variety of workshops.

A key component of teaching is mentoring students and postdocs. I've mentored four graduate students and fifteen postdocs directly. See [here](people.qmd) for more information.


## Service


* I served as an associate head for the EEB department for several years, focusing on graduate students, plus additional duties like helping to craft a covid response plan. In that role I worked to address student concerns and try to improve climate, and I continued that work as a member of EEB's diversity committee. 
* I was also the associate director for postdoctoral activities at the National Institute for Biological and Mathematical Synthesis (NIMBioS), managing training and evaluation of postdocs. Other service includes organizing things like hackathons to draw in new participants and workshops to train biologists in new tools. 
* I also co-organized the iEvoBio conference for three years and helped with other biology conferences (for example, scheduling all the lightning talks at a couple of Evolution meetings). 
* I've done workshops to help with students applying to grants, overseen grad admissions, etc. 
* I have been heavily involved in the Society of Systematic Biologists, in the past serving as the communications director, a member of the council, and its president. I have also been one of the longest serving members the joint code of conduct committee for the Evolution meetings, where have been part of the team creating [Safe Evolution](https://www.evolutionmeetings.org/safe-evolution.html), as well as doing IRB-approved research on climate at our meetings ([poster](https://figshare.com/articles/poster/Unwelcome_behaviors_at_the_Evolution_meetings_Survey_results/8311595)). 
* I have served as faculty advisor for Darwin Day Tennessee, one of the oldest Darwin Day events in the US (our activities now range from teacher workshops to bringing in speakers like Camile Parmesan and Neil Shubin to having giant puppets of Darwin and Wallace greet children at a birthday party for Darwin. 
* I also try to make biology a better, more diverse place, whether it's through helping sustain our department's women in science group to talking about imposter syndrome with graduate students to finding ways to get a diverse set of applicants for our graduate program to creating a [tool](better-letter.qmd) to allow people to self-evaluate letters of rec they are writing for words associated with biases.


## Funding

In the last ten years, work in the lab has been funded by six NSF grants to me as PI or Co-PI as well as awards from iPlant, Encyclopedia of Life, and Google Summer of Code to me or people doing work in the lab (see more info [here](funding.qmd)). Grad students in the lab have been supported by teaching assistantships as well as a [PEER](https://web.utk.edu/~peer/) fellowship. The UT Knoxville-based [National Institute for Mathematical and Biological Synthesis (NIMBioS)](https://www.nimbios.org/) remains critically important for my work, whether by funding independent postdocs (I have mentored multiple NIMBioS postdocs, in addition to additional postdocs in my lab with other funds), sponsoring workshops, or organizing working groups.

## Contact

To contact me, bomeara@utk.edu. I am also on <a rel="me" href="https://mastodon.social/@omearabrian">Mastodon</a> and [bluesky](https://bsky.app/profile/omearabrian.bsky.social). Office hours, including virtual ones are at my [calendar](calendar.qmd).